---
title: "Music Insights"
author: "John Higdon"
output: html_notebook
---

First, we need tidyverse to handle data importing and wrangling...

```{r}
install.packages("tidyverse")
library("tidyverse")

survey <- read_csv("https://raw.githubusercontent.com/introdsci/MusicSurvey/master/music-survey.csv")
preferences <- read_csv("https://raw.githubusercontent.com/introdsci/MusicSurvey/master/preferences-survey.csv")

colnames(survey)[colnames(survey) == "What is your pseudonym?"] <- "Pseudonym"
```

Multiple classmates let me know what we did in the class of 9/26, which I missed due to being called into work (not an excused absence, I let my employer know how upset I was about this.).  

Let's make our dataframes more readable:

```{r}
colnames(survey)[colnames(survey) == "Year you were born (YYYY)"] <- "year_born"
colnames(survey)[colnames(survey) == "Timestamp"] <- "date_submitted"
colnames(survey)[colnames(survey) == "Which musical instruments/talents do you play? (Select all that apply)"] <- "instruments_played"
colnames(survey)[colnames(survey) == "Academic Year"] <- "academic_year"

# using column numbers to enact changes
colnames(survey)[2] <- "psuedonym_generated"
colnames(survey)[9] <- "favorite_artist"
colnames(survey)[11] <- "favorite_song_url"
colnames(survey)[10] <- "favorite_song"

```

Now let's load dplyr and tidyr:
```{r}
library("dplyr")
library("tidyr")
```

Let's create new tables using tibble.  
We're going to build a table that shows information about the people who responded.
```{r}
person_table <- tibble(date_submitted = survey$date_submitted, pseudonym = survey$Pseudonym, sex = survey$Sex, academic_year = survey$academic_year, major = survey$Major, year_born = survey$year_born, pseudonym_generated = survey$psuedonym_generated)
```

Finally, let's create a new table with favorite songs.  
```{r}
fav_songs <- tibble(pseudonym = survey$Pseudonym, song = survey$favorite_song, artist = survey$favorite_artist, link = survey$favorite_song_url)
```

Use parse_datetime to determine how characters of the date and time are defined within R survey results
```{r}
person_table$date_submitted <- as.POSIXlt(parse_datetime(person_table$date_submitted, format="%m/%d/%y %H:%M"))
```
