---
title: "Music Insights"
author: "John Higdon"
output: html_notebook
---

First, we need tidyverse to handle data importing and wrangling...

```{r}
#install.packages("tidyverse")
library("tidyverse")

survey <- read_csv("https://raw.githubusercontent.com/introdsci/MusicSurvey/master/music-survey.csv")
preferences <- read_csv("https://raw.githubusercontent.com/introdsci/MusicSurvey/master/preferences-survey.csv")

colnames(survey)[colnames(survey) == "What is your pseudonym?"] <- "Pseudonym"
```

Let's make our dataframes more readable:

```{r}
colnames(survey)[colnames(survey) == "Year you were born (YYYY)"] <- "year_born"
colnames(survey)[colnames(survey) == "Timestamp"] <- "date_submitted"
colnames(survey)[colnames(survey) == "Which musical instruments/talents do you play? (Select all that apply)"] <- "instruments_played"
colnames(survey)[colnames(survey) == "Academic Year"] <- "academic_year"

# using column numbers to enact changes
colnames(survey)[2] <- "psuedonym_generated"
colnames(survey)[9] <- "favorite_artist"
colnames(survey)[11] <- "favorite_song_url"
colnames(survey)[10] <- "favorite_song"
colnames(preferences)[2] <- "pseudonym"
```

Now let's load dplyr and tidyr:
```{r}
library("dplyr")
library("tidyr")
```

Let's create new tables using tibble.  
We're going to build a table that shows information about the people who responded.
```{r}
person_table <- tibble(date_submitted = survey$date_submitted, pseudonym = survey$Pseudonym, sex = survey$Sex, academic_year = survey$academic_year, major = survey$Major, year_born = survey$year_born, pseudonym_generated = survey$psuedonym_generated)
```

Finally, let's create a new table with favorite songs.  
```{r}
fav_songs <- tibble(pseudonym = survey$Pseudonym, song = survey$favorite_song, artist = survey$favorite_artist, link = survey$favorite_song_url)
```

Use parse_datetime to determine how characters of the date and time are defined within R survey results
```{r}
person_table$date_submitted <- as.POSIXlt(parse_datetime(person_table$date_submitted, format="%m/%d/%y %H:%M"))
```

Convert columns to factors as needed
```{r}
person_table$academic_year <- as.factor(person_table$academic_year)
person_table$major <- as.factor(person_table$major)
```

Change level to be more consistent
```{r}
levels(person_table$major)[levels(person_table$major) == "Computer information systems"] <- "Computer Information Systems"
```

Create Ratings table using gather
```{r}
ratings <- tidyr::gather(preferences, artist_song, rating, 3:45)
```

Now let's create some visulizations of our data.
```{r}
library(ggplot2)

# adds colors for each user that submitted a vote
ggplot(ratings, aes(x = rating, fill = pseudonym)) + geom_histogram(binwidth = 1)

ggplot(ratings, aes(x = rating, y = pseudonym)) + geom_boxplot()  # there's a problem here. Not able to view boxplot.
ggplot(ratings, aes(x = pseudonym, y = rating)) + geom_boxplot()

# Box plot with differing colors per pseudonym used
ggplot(ratings, aes(x = pseudonym, y = rating, fill = pseudonym)) + geom_boxplot()

# Clean up.
ggplot(ratings, aes(x = pseudonym, y = rating, fill = pseudonym)) + geom_boxplot() + theme_light() + coord_flip()
```

The following is sourced from Professor Buffardi's repo that was presented 10/25. Source: https://github.com/introdsci/MusicInsights/blob/ref/insights.Rmd

The given code cleans up responses to allow for easier cateogrizing so we can better tidy our data.

```{r}
MusicalTalents <- tibble(pseudonym = survey$Pseudonym, instruments_played = survey$instruments_played)

MusicalTalents <- separate_rows(MusicalTalents,-pseudonym, sep=",")
#clean responses for consistency:
MusicalTalents$instruments_played <- MusicalTalents$instruments_played %>% 
                                     trimws() %>% 
                                     tolower() %>%
                                     str_replace_all(pattern=".*piano.*","piano") %>%
                                     str_replace_all(pattern=".*ukulele.*","ukulele") %>%
                                     as.factor()
```

Mistakes were made by two users. Let's filter out their results. Source is from Professor Buffardi's repo mentioned above.

```{r}
earliest_time <- min(ratings$Timestamp[ratings$pseudonym=="Angel Angel"])
ratings_new <- ratings %>% filter(!(pseudonym=="Angel Angel" & Timestamp!=earliest_time))
earliest_time <- min(ratings$Timestamp[ratings$pseudonym=="Mission Theory"])
ratings_new <- ratings %>% filter(!(pseudonym=="Mission Theory" & Timestamp!=earliest_time))

ratings_new$artist_song <- str_replace_all(ratings_new$artist_song, 
                           pattern = "\t", 
                           replacement = " ")
```

Now we can view the favorite song of each person who participated in the survey.

```{r}
favorite_rating <- ratings_new %>% 
    left_join(fav_songs, by="pseudonym") %>%
    filter(artist_song==paste(artist,song)) %>%
    select(pseudonym,artist_song,rating)
```

